Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-08-09 surfaces-write-up/surface-write-up.tex
minor rewording, add one clarification todo

\noindent
In section \ref{sec:hirz}, we shift our attention to Hirzebruch Surfaces and prove the following theorem:
\begin{restatable}{thm}{hirzrestate}
\label{thm:hirz-generators-relations}
Let $u,v,z,w$ be the coordinates for the Hirzebruch surface $\hirz_m$, as described at the beginning of section ~\ref{sec:hirz}. Let $D = \sum_{i=1}^n \alpha_i D_i \in \bq \otimes_\bq \di \hirz_m$ 
where $\alpha_i = \frac{c_i}{k_i}$.
Then notate
\[\ell_{i,j} = \lcm_{h\ne i,j}(k_h).\] 
Let $D_i =
V(f_i)$ where $f_i \in \sco(a_i, b_i)$; further suppose $f_1$ and $f_2$ are
independent linear
polynomials in $u$ and $v$ and $f_3$ and $f_4$ are independent
linear polynomials in $w$ and $x$.
Then $R_D$ is generated in degrees at most
\[
	\rho = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i +
	\sum_{\substack{i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j
	- \bida_j \bidb_i) \ell_{i, j}
\]

@@

Our first aim in this section is to prove Theorem
~\ref{thm:proj-generators-relations} bounding the number of generators and
relations of $R_D$. Before doing so, we justify the importance of assumption
~\eqref{eqn:pm-basis-assumption} with several illustrative examples.

@@

For the remainder of this section we will assume $D_1, D_2, D_3,$ and $D_4$ are
distinct divisors with bi-degrees $(1,0)$
, $(1,0)$, $(0,1)$, and $(0,1)$ respectively with $D_i = V(f_i)$ for $1 \leq i \leq 4$ with $f_i \in \sco(\bida_i, \bidb_i)$.  
In order to achieve this condition on
$D_1, \ldots D_4$, it may be necessary to add in ``ghost divisors'' (i.e.
divisors with of the desired form with a coefficient $0$).
Also, note that, $f_1$ and $f_2$ are independent linear
polynomials in $u$ and $v$ and $f_3$ and $f_4$ are independent
linear polynomials in $z$ and $w$.
Analogously to Proposition ~\ref{prop:pm-span-and-basis} for the case
of $\bp^m$, all rational functions on $\hirz_m$
can be written uniquely in a form where their numerator is a function
of only $f_1,f_2,f_3$, and $f_4$.
\todo{Aaron: Do either of you think we should write out this as
a proposition more clearly? I'm fine either way, but it was something
missing which I thought I should point out. \\
Peter: The one line argument is ``Every numerator is a polynomial in $u, v, z, w$; we can generate $u$ and $v$ from $f_1$ and $f_2$, and we can generate $z$ and $w$ from $f_3$ and $f_4$.''  This can be added if you want

>>>>

\noindent
In Section \ref{sec:hirz}, we shift our attention to Hirzebruch Surfaces and prove the following theorem:
\begin{restatable}{thm}{hirzrestate}
\label{thm:hirz-generators-relations}
Let $u,v,z,w$ be the coordinates for the Hirzebruch surface $\hirz_m$, as described at the beginning of Section ~\ref{sec:hirz}. Let $D = \sum_{i=1}^n \alpha_i D_i \in \bq \otimes_\bq \di \hirz_m$ 
where $\alpha_i = \frac{c_i}{k_i}$.
Then notate
\[\ell_{i,j} = \lcm_{h\ne i,j}(k_h).\] 
Let $D_i =
V(f_i)$ where $f_i \in \sco(a_i, b_i)$; further suppose $f_1$ and $f_2$ are
independent linear
polynomials in $u$ and $v$ and $f_3$ and $f_4$ are independent
linear polynomials in $w$ and $x$.
Then $R_D$ is generated in degrees at most
\[
	\rho = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i +
	\sum_{\substack{i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j
	- \bida_j \bidb_i) \ell_{i, j}
\]

@@

The first aim of this section is to prove Theorem
~\ref{thm:proj-generators-relations} bounding the number of generators and
relations of $R_D$. Before doing so, we justify the importance of assumption
~\eqref{eqn:pm-basis-assumption} with several illustrative examples.

@@

For the remainder of this section, we will assume that $D_1, D_2, D_
3, D_4$ are distinct divisors with bi-degrees $(1, 0)$, $(1, 0)$, $(
0, 1), (0, 1)$ respectively with $D_i = V(f_i)$ for $1 \leq i \leq 4
$ with $f_i \in \sco(\bida_i, \bidb_i)$.

In order to achieve this condition on \todo{write ``this''}
$D_1, \ldots D_4$, it may be necessary to add in ``ghost divisors'' (i.e.
divisors with of the desired form with a coefficient $0$).
Also, note that, $f_1$ and $f_2$ are independent linear
polynomials in $u$ and $v$ and $f_3$ and $f_4$ are independent
linear polynomials in $z$ and $w$.
Analogously to Proposition ~\ref{prop:pm-span-and-basis} for the case
of $\bp^m$, all rational functions on $\hirz_m$
can be written uniquely in a form where their numerator is a function
of only $f_1,f_2,f_3$, and $f_4$.
\todo{Aaron: Do either of you think we should write out this as
a proposition more clearly? I'm fine either way, but it was something
missing which I thought I should point out. \\
Peter: The one line argument is ``Every numerator is a polynomial in $u, v, z, w$; we can generate $u$ and $v$ from $f_1$ and $f_2$, and we can generate $z$ and $w$ from $f_3$ and $f_4$.''  This can be added if you want
